---
title: "W6_Loops_Functions"
author: "Jakub Raszka"
date: "26 10 2020"
output:
  word_document: default
  html_document:
    df_print: paged
  pdf_document: default
---

## Loading data, packages
```{r}
## packages
pacman::p_load(ggplot2, tidyverse, gapminder, psych)

## data
gapminder <- gapminder


```


## 1: Function

Define a defensive function that calculates the Gross Domestic Product of a nation from the data available in the gapminder dataset. Using that function, calculate the GDP of Denmark in the following years: 1967, 1977, 1987, 1997, 2007, and 2017.
```{r}
## defining function
gdp_calculator <- function(nation, year, population, gdp_per_capita){
  GDP <- population * gdp_per_capita
  
  return(GDP)
}

## creating a dataframe of data only from Denmark and for desired years
denmark_7 <- gapminder %>% 
  filter(country == "Denmark") %>% 
  filter(grepl("\\d{2}[1,6,7,8,9,0]7", year))

## upon close inspection, there is no data for year 2017!!

## calculating gdp of country
denmark_7$gdp_country <- gdp_calculator(denmark_7$country, denmark_7$year, denmark_7$pop, denmark_7$gdpPercap)

denmark_7
```


## 2: Loop

Write a script that loops over each country in the gapminder dataset, tests whether the country starts with a ‘B’ , and print out whether the life expectancy is smaller than 50, between 50 and 70, or greater than 70.

```{r}
## a loop prints whether a country name start with B or not, but what's the point??

# creating a vector with unique country name (to prevent repetition)
country_name <- sort(unique(as.character(gapminder$country)), decreasing =F)

# a loop 
## to prevent loops output to be too long, I restricted the conditon only to first 20 countries. However, it can be easily undone by deleting [1:20] after country name
for (i in 1:length(country_name[1:20])){
  
  if(grepl("^B", country_name[i]) == TRUE){
  print(paste(country_name[i], " starts with B."))
  } else (print(paste(country_name[i], " does not start with B.")))

}

## creating an empty vector to store results
gapminder$lifeexp_task <- c("l")


## printing a message in what range life expectancy is

for (i in 1:nrow(gapminder)){
  
  if(gapminder$lifeExp[i]< 50){
  
    gapminder$lifeexp_task[i] <- "The life expectancy is smaller than 50"
    
  } else if (gapminder$lifeExp[i]>= 50 & gapminder$lifeExp[i] <= 70){
   
    gapminder$lifeexp_task[i] <- "The life expectancy is between  50 and 70."
  
    } else(gapminder$lifeexp_task[i] <- "The life expectancy is larger than 70.")

}

gapminder[10:20,c(4,7)]
```




